\documentclass{llncs}
\usepackage{epsfig,graphics,latexsym}
\usepackage{amsmath,amssymb,theorem,enumerate}
\usepackage[draft]{fixme}


%% Farbige Kommentar-Felder
\usepackage[usenames]{color} % Only used in comment commands


\newcommand{\commentout}[1]{}



\definecolor{Blue}{rgb}{0,0.16,0.90}
\definecolor{Red}{rgb}{0.90,0.16,0}
\definecolor{DarkBlue}{rgb}{0,0.08,0.45}
\definecolor{ChangedColor}{rgb}{0.9,0.08,0}
\definecolor{CommentColor}{rgb}{0.2,0.8,0.2}
\definecolor{ToDoColor}{rgb}{0.1,0.2,1}

\definecolor{verylightgray}{rgb}{0.91,0.91,0.91}

% *** Use this definition of the command to show the comments ***
\newcommand{\todo}[1]{\textbf{\color{ToDoColor} TODO: #1}}
\newcommand{\joerg}[1]{\textbf{\color{CommentColor} /* #1  (joerg)*/}}
\newcommand{\alex}[1]{\textbf{\color{CommentColor} /* #1  (alex)*/}}
\newcommand{\carlos}[1]{\textbf{\color{CommentColor} /* #1  (carlos)*/}}




%\newenvironment{proof}{\noindent {\bf Proof:}}%
%{\hfill \rule[0.3ex]{1ex}{1ex} \par \addvspace{\bigskipamount}}

\newenvironment{proofsketch}{\noindent {\bf Proof Sketch:}}%
{\hfill \rule[0.3ex]{1ex}{1ex} \par \addvspace{\bigskipamount}}




\title{When Operators Learned How to Split Up}

\author{Koller, Areces, Hoffmann, \& Friends}


\institute{I'm sure they're working somewhere \dots}






\begin{document}


\maketitle


\begin{abstract}
PDDL is arguably a programming language for planners. While that
language, in the case of ``fully automatic'' planners, is limited to a
description of the mechanics of the domain of interest, it is
well-known that different models of the same domain can yield immense
differences in planner performance. Indeed, to a considerable extent
the domain-designing activities of hitherto IPC organizers consisted
of ``spoon-feeding'' the targeted domains to the planners, in that
they used their inside knowledge of planning algorithms to design the
domains in a way so that existing planners could attack them in a
sensible way. However, what if actual ``end-users'', people that want
to use planners but that haven't spent the last 10 years implementing
them, wish to ``spoon-feed'' their domain to the planners?

To some extent, this can be supported by documentation and best
practices explanations, as well as graphical environments for domain
modeling. However, the holy grail would be to provide
\emph{domain-rewriting} techniques, in a manner similar to what is
known as program re-writing: Given an input model designed by the
end-user, can we automatically modify that model to yield better
performance, while still encoding essentially the same domain?

We herein begin to address this question, focussing on one particular
technical detail that affects the pre-processing (and in certain ways
also the search techniques) of many existing planners: the number of
operator parameters. For example, in the development of the Pipesworld
domain as used at IPC 2004, one of the essential steps was to split up
some of the operators, so that each part would have less parameters
and thus be more feasible for common pre-processing techniques. We
herein provide an automatic method that does just that. We evaluate
the method on IPC benchmarks, as well as on the domain of sentence
generation, where pre-processing, in particular related to the number
of operator parameters, is known to be one of the major obstacles.
\end{abstract}




\setcounter{tocdepth}{4}

\tableofcontents





\section{Operator Splitting}

\input{opsplit-definition-v1}

\input{opsplit-definition-v2}




\section{Task/Domain Equivalence}

\input{equiv-definition-carlos}

\input{equiv-definition-alex}

\input{equiv-definition-gabi}




\bibliographystyle{plain}
\bibliography{biblio}


\end{document}
