\section{Introduction}
\label{sec:introduction}

\emph{Natural language generation} (NLG; \citealp{reiter00building}) is one of
the major subfields of natural language processing, concerned with computing
natural language sentences or texts that convey a given piece of information to
an audience. While the output of a generation task can take many forms,
including written text, synthesised speech, or embodied multimodal
presentations, the underlying NLG problem in each case can be modelled as a
problem of achieving a (communicative) goal by successively applying a set of
(communicative) actions. This view of NLG as goal-directed action has clear
parallels to \emph{automated planning}, which seeks to find general techniques
for efficiently solving the action sequencing problem.

Treating generation as planning has a long history in NLG, ranging
from the initial attempts of the field to utilise early planning
approaches \citep{perrault80,appelt:planning,hovy88,young94dpocl}, to
a recent surge of research
\citep{Steedman-Petrick:07,KolSto07,brenner08:_contin_multiag_plann_approac_to_situat_dialog,benotti08b}
seeking to capitalise on the improvements modern planners offer in
terms of efficiency and expressiveness. This paper attempts to assess
the usefulness of current planning techniques to NLG by investigating
some representative generation problems, and evaluating whether
automated planning has advanced to the point that it can provide
solutions to such NLG applications---applications that are not
currently being investigated by mainstream planning research.

To answer this question, we proceed in two steps. First, we present
two generation problems that have recently been cast as planning
problems: the sentence generation task and the GIVE task. In the
sentence generation task, we concentrate on generating a single
sentence that expresses a given meaning. In this case, a plan encodes
the necessary sentence with the actions in the plan corresponding to
the utterance of individual words \citep{KolSto07}. In the GIVE domain
(``Generating Instructions in Virtual Environments''), we describe a
new shared task that was recently posed as a challenge for the NLG
community \citep{ByrKolStrCasDalMooObe09}. GIVE uses planning as part
of a larger NLG system for generating natural-language instructions
that guide a human user in performing a given task in a virtual
environment.

Second, we evaluate the performance of current off-the-shelf planners
in the planning domains into which these two generation problems
translate. Among the planners we test, we explore the efficiency of FF
\citep{HoffmannNebel01}---a planner that has arguably had the greatest
impact on recent approaches to deterministic planning---and some of its
descendants, such as SGPLAN
\citep{hsu06:_new_featur_in_sgplan_for}. All of the planners we test
are freely available, support an expressive subset of the Planning
Domain Definition Language (PDDL; McDermott et al.~\citeyear{PDDL}),
and have been successful on both standard planning benchmarks and the
problems of the International Planning Competition (IPC).\footnote{See
  \url{http://ipc.icaps-conference.org/} for information about recent
  and past editions of the IPC. A good overview of the deterministic
  track of the 2004 instance of the IPC can be found in
  \citep{Hoffmann-Edelkamp:05}.}  Using these planners---and an ad-hoc
Java implementation of GraphPlan \citep{Blum1997} which serves as a
baseline for certain experiments---we perform a series of tests on a
range of problem instances in our NLG domains.

Overall, our findings are mixed. On the one hand, we demonstrate that
some planners can readily handle the \emph{search} problems that arise
in our testing domains on realistic inputs, which is promising given
the challenging nature of these tasks (e.g., the sentence generation
task is NP-complete; see \citealt{KolStr02}). On the other hand, these
same planners spend tremendous amounts of time on \emph{preprocessing}
to analyse the problem domain in support of the search. On many of our
problem instances, the preprocessing time overshadows the search time.
For instance, FF spends 90\% of its runtime in the sentence generation
domain on grounding out actions and literals. As a consequence, we
conclude that the off-the-shelf planners we investigated are too slow
to be useful in real NLG applications.  We hope that our results will
spark some interest for more efficient implementations of planning
preprocessors, and offer our NLG domains and experiences as challenges
for the planning community.

The remainder of this paper is structured as follows. In
Section~\ref{sec:nlg-as-planning}, we introduce the idea of NLG as
planning and briefly review the relevant literature. In
Section~\ref{sec:domains}, we describe a set of planning problems
associated with two NLG tasks, namely sentence planning and situated
instruction generation. In Section~\ref{sec:experiments}, we report on
the results of our experiments with these planning problems.  In
Section~\ref{sec:discussion} we discuss our results and overall
experiences, and conclude in Section~\ref{sec:conclusion}.


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "manuscript"
%%% End: 
